Performing Cluster Analysis on TAD boundaries & Annotations

### 2019-03-22 Scripts/evaluating_pca.Rmd
editted pca script

tfbsvars <-  gsub(".bed","",list.files("Z:/TAD_data_analysis/annotations/tfbs/cell_type_specific/gm12878"))

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_count_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_perc_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_dist_tfbs,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_tfbs, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

>>>>

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_count_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_count_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_count_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_count_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_perc_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_perc_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_perc_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_perc_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

@@

##with loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_10kb_dist_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_10kb_dist_tfbs)-1)*sum(apply(pcadat_gm12878_10kb_dist_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_10kb_dist_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_10kb_dist_tfbs, 5, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_10kb_dist_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_count.rds")
pcadat_gm12878_25kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_perc.rds")
pcadat_gm12878_25kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/25kb/Ensemble/PCA/pcadat_gm12878_25kb_dist.rds")

#performing log2 transformation on distances
pcadat_gm12878_25kb_dist <- apply(pcadat_gm12878_25kb_dist,2,function(x){log(x + 1, base=2)})

colnames(pcadat_gm12878_25kb_count) <- cleancolnames(pcadat_gm12878_25kb_count)
colnames(pcadat_gm12878_25kb_perc) <- cleancolnames(pcadat_gm12878_25kb_perc)
colnames(pcadat_gm12878_25kb_dist) <- cleancolnames(pcadat_gm12878_25kb_dist)

pcadat_gm12878_25kb_count <- t(pcadat_gm12878_25kb_count)
pcadat_gm12878_25kb_perc <- t(pcadat_gm12878_25kb_perc)
pcadat_gm12878_25kb_dist <- t(pcadat_gm12878_25kb_dist)

#pcadat_gm12878_25kb_count <- data.frame(pcadat_gm12878_25kb_count)
#pcadat_gm12878_25kb_perc <- data.frame(pcadat_gm12878_25kb_perc)
#pcadat_gm12878_25kb_dist <- data.frame(pcadat_gm12878_25kb_dist)

#pcadat_gm12878_25kb_count$Type <- "Counts"
#pcadat_gm12878_25kb_perc$Type <- "Percent"
#pcadat_gm12878_25kb_dist$Type <- "Distance"

```{r}
pcadat_gm12878_25kb_count_chromstates <- pcadat_gm12878_25kb_count[which(rownames(pcadat_gm12878_25kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_count_chromstates)-1)*sum(apply(pcadat_gm12878_25kb_count_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_count_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_count_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_count_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_perc_chromstates <- pcadat_gm12878_25kb_perc[which(rownames(pcadat_gm12878_25kb_perc) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_perc_chromstates)-1)*sum(apply(pcadat_gm12878_25kb_perc_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_perc_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_perc_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_perc_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_dist_chromstates <- pcadat_gm12878_25kb_dist[which(rownames(pcadat_gm12878_25kb_dist) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_dist_chromstates)-1)*sum(apply(pcadat_gm12878_25kb_dist_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_dist_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_dist_chromstates, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_dist_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_count_histones <- pcadat_gm12878_25kb_count[which(rownames(pcadat_gm12878_25kb_count) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_count_histones)-1)*sum(apply(pcadat_gm12878_25kb_count_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_count_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_count_histones, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_count_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_perc_histones <- pcadat_gm12878_25kb_perc[which(rownames(pcadat_gm12878_25kb_perc) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_perc_histones)-1)*sum(apply(pcadat_gm12878_25kb_perc_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_perc_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_perc_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_perc_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_dist_histones <- pcadat_gm12878_25kb_dist[which(rownames(pcadat_gm12878_25kb_dist) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_dist_histones)-1)*sum(apply(pcadat_gm12878_25kb_dist_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_dist_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_dist_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_dist_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_count_tfbs <- pcadat_gm12878_25kb_count[which(rownames(pcadat_gm12878_25kb_count) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_count_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_count_tfbs)-1)*sum(apply(pcadat_gm12878_25kb_count_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_count_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_count_tfbs, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_count_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_perc_tfbs <- pcadat_gm12878_25kb_perc[which(rownames(pcadat_gm12878_25kb_perc) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_perc_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_perc_tfbs)-1)*sum(apply(pcadat_gm12878_25kb_perc_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_perc_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_perc_tfbs, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_perc_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_25kb_dist_tfbs <- pcadat_gm12878_25kb_dist[which(rownames(pcadat_gm12878_25kb_dist) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_25kb_dist_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_25kb_dist_tfbs)-1)*sum(apply(pcadat_gm12878_25kb_dist_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_25kb_dist_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_25kb_dist_tfbs, 5, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_25kb_dist_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_count <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_count.rds")
pcadat_gm12878_50kb_perc <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_perc.rds")
pcadat_gm12878_50kb_dist <- readRDS("Z:/TAD_data_analysis/GM12878/50kb/Ensemble/PCA/pcadat_gm12878_50kb_dist.rds")

#performing log2 transformation on distances
pcadat_gm12878_50kb_dist <- apply(pcadat_gm12878_50kb_dist,2,function(x){log(x + 1, base=2)})

colnames(pcadat_gm12878_50kb_count) <- cleancolnames(pcadat_gm12878_50kb_count)
colnames(pcadat_gm12878_50kb_perc) <- cleancolnames(pcadat_gm12878_50kb_perc)
colnames(pcadat_gm12878_50kb_dist) <- cleancolnames(pcadat_gm12878_50kb_dist)

pcadat_gm12878_50kb_count <- t(pcadat_gm12878_50kb_count)
pcadat_gm12878_50kb_perc <- t(pcadat_gm12878_50kb_perc)
pcadat_gm12878_50kb_dist <- t(pcadat_gm12878_50kb_dist)

#pcadat_gm12878_50kb_count <- data.frame(pcadat_gm12878_50kb_count)
#pcadat_gm12878_50kb_perc <- data.frame(pcadat_gm12878_50kb_perc)
#pcadat_gm12878_50kb_dist <- data.frame(pcadat_gm12878_50kb_dist)

#pcadat_gm12878_50kb_count$Type <- "Counts"
#pcadat_gm12878_50kb_perc$Type <- "Percent"
#pcadat_gm12878_50kb_dist$Type <- "Distance"

```{r}
pcadat_gm12878_50kb_count_chromstates <- pcadat_gm12878_50kb_count[which(rownames(pcadat_gm12878_50kb_count) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_count_chromstates)-1)*sum(apply(pcadat_gm12878_50kb_count_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_count_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_count_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_count_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_perc_chromstates <- pcadat_gm12878_50kb_perc[which(rownames(pcadat_gm12878_50kb_perc) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_perc_chromstates)-1)*sum(apply(pcadat_gm12878_50kb_perc_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_perc_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_perc_chromstates, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_perc_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_dist_chromstates <- pcadat_gm12878_50kb_dist[which(rownames(pcadat_gm12878_50kb_dist) %in% chromstatevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_chromstates),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_chromstates),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_dist_chromstates)-1)*sum(apply(pcadat_gm12878_50kb_dist_chromstates,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_dist_chromstates,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_dist_chromstates, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_dist_chromstates,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_count_histones <- pcadat_gm12878_50kb_count[which(rownames(pcadat_gm12878_50kb_count) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_count_histones)-1)*sum(apply(pcadat_gm12878_50kb_count_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_count_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_count_histones, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_count_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_perc_histones <- pcadat_gm12878_50kb_perc[which(rownames(pcadat_gm12878_50kb_perc) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_perc_histones)-1)*sum(apply(pcadat_gm12878_50kb_perc_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_perc_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_perc_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_perc_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_dist_histones <- pcadat_gm12878_50kb_dist[which(rownames(pcadat_gm12878_50kb_dist) %in% histonevars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_histones),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_dist_histones),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_dist_histones)-1)*sum(apply(pcadat_gm12878_50kb_dist_histones,2,var))
  for (i in 2:5) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_dist_histones,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:5, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_dist_histones, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_dist_histones,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_count_tfbs <- pcadat_gm12878_50kb_count[which(rownames(pcadat_gm12878_50kb_count) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_count_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_count_tfbs)-1)*sum(apply(pcadat_gm12878_50kb_count_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_count_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_count_tfbs, 3, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_count_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_perc_tfbs <- pcadat_gm12878_50kb_perc[which(rownames(pcadat_gm12878_50kb_perc) %in% tfbsvars),]

##with loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_tfbs),
         loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##without loadings
autoplot(prcomp(pcadat_gm12878_50kb_perc_tfbs),
         #loadings = TRUE,
         label = TRUE, 
         label.size = 3,
         label.repel=T) +
  theme_bw()
##kmeans clustering algorithm to group annotations
wss <- (nrow(pcadat_gm12878_50kb_perc_tfbs)-1)*sum(apply(pcadat_gm12878_50kb_perc_tfbs,2,var))
  for (i in 2:50) wss[i] <- sum(kmeans(pcadat_gm12878_50kb_perc_tfbs,
                                       centers=i, nstart=25, iter.max = 1000)$withinss)
plot(1:50, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")

set.seed(1)
autoplot(kmeans(pcadat_gm12878_50kb_perc_tfbs, 4, nstart=25, iter.max = 1000),
         data=pcadat_gm12878_50kb_perc_tfbs,
         #frame = TRUE,
         #frame.type = 'norm',
         label = TRUE, 
         label.size = 3, 
         label.repel=T,
         label.show.legend = FALSE) + 
  #guides(fill=FALSE) +
  theme_bw()

```{r}
pcadat_gm12878_50kb_dist_tfbs <- pcadat_gm12878_50kb_dist[which(rownames(pcadat_gm12878_50kb_dist) %in% tfbsvars),]
